Gap formation and stability in non-isothermal protoplanetary discs

### 2014-11-19 paper/results.tex
Updated figures

\subsection{Generic evolution} 
The linear growth of the RWI and vortex-formation is followed by 
vortex merging. We now find merging timescales independent of
$\tilde\beta$, and by $60P_o$ only one vortex remains.  
%as the growth of these modes are now 
%accelerated by planetary influence as apposed to `planet off' simulations.
% i.e. only the $m=1$ mode remains. 
The evolution of the amplitude of the $m=1$ surface density component,
averaged over $r-r_p\in[2,10]r_h$ is shown in Fig~\ref{lifetimeplot} 
for different $\tilde\beta$. 
{\bf any particular reason for the larger radial range for average?
  did we get bigger vortices?} 
% {\bf should state (also for planet-off runs)
% what the radial range is used for averaging the Fourier amplitudes}

@@

In all cases the system remains in a quasi-steady state for
$\gtrsim800P_0$ with a single vortex circulating 
the outer gap edge at the local Keplerian  
frequency. Fig. \ref{Vortex2D} shows a typical plot of the relative 
surface density perturbation in this state. During this stage, the 
vortex intensifies. As the vortex forms it has an initial 
Rossby number $Ro\approx-0.1$ for all $\tilde\beta$
(indicating anti-cyclonic
motion). As the $m=1$ amplitude grows the Rossby number also increases
in magnitude, 
%relative spin of the vortex
%continuously grows along with it
, reaching a characteristic value of $Ro\approx-0.4$ for all cooling
rates. {\bf is the the value reached just before decay for all cooling?}

@@

We also measured large increases in the Mach number near
the vortex when it begins to decay. Figure \ref{machplot} plots the
Mach number $M=|\bm{v} - \bm{v}_\mathrm{vor}|/c_s$, where
$\bm{v}_\mathrm{vor}$ corresponds to the bulk velocity of the vortex
around the disc. {\bf check - i assume this is the definition of mach
  number used here}. Values in Fig. \ref{machplot} have been averaged over
a region within $2H$ of the vortex centre.    
% Mach numbers with repect to the vortex bulk motion around the vortex
% region of $|r-r_{vortex}|<2H$
% are shown in  for various cooling rates over time.
During the quasi-steady state the Mach number increase
steadily, and for all cases $M$ reaches a maximum at 
% with larger growth rate for higher $\tilde\beta$.
a time coincident with the start of vortex decay.

@@

\subsection{Effect of vortex dissipation on the gap structure} 
{\bf are these high resolution results?}
We find vortex growth during the quasi-steady state and its dissipation
also modifies the gap structure. Here, we 
define a gap depth parameter by taking the azimuthal and radial average
of the relative surface density perturbation within $r\in[r_p,r_e]$ where
$r_{e}$ is defined as the outer gap radius such that 
$\langle\Delta \Sigma(r_e)/\Sigma(t=0)\rangle_{\phi}=0$. We are
interested in the region $r \in[r_p,r_e]$ since the vortex is
located at the outer gap edge. A more negative gap depth parameter
correlates with a steeper gap edge. 
{\bf notation clash: rout already used as outer disc boundary - choose
different notation. updated to $r_e$ for now}

@@

@@

These results were reproduced in lower resolution simulations as well. Vortex
shocks were similarly seen during vortex dissipation and non-monotonic
 dependence on lifetime with $\tilde\beta$ was similarly found.
Vortex lifetimes are considerably longer in these runs with the maximum
 lifetime longer than the simulation timescale of $2000P_o$ for 
$\tilde\beta=5.0$ which again has $h\approx0.06$.
This inconsistancy in timescales, may be due to resolution as a finer 
grid spacing is required to accurately
reproduce jump conditions along shocks and which cause vortex dissipation.
 {\bf this is the best reason I could personally think of, for lower
   beta runs vortices last slightly longer but for higher values it is
   quite drasticly different. Not sure how numerical dissipation as
   you mentioned would be a factor in this case}

>>>>

\subsection{Generic evolution} 
The linear growth of the RWI and vortex-formation is followed by 
vortex merging. We now find merging timescales independent of
$\tilde\beta$, and by $60P_o$ only one vortex remains.  
%as the growth of these modes are now 
%accelerated by planetary influence as apposed to `planet off' simulations.
% i.e. only the $m=1$ mode remains. 
The evolution of the amplitude of the $m=1$ surface density component,
averaged over $r-r_p\in[2,10]r_h$ is shown in Fig~\ref{lifetimeplot} 
for different $\tilde\beta$. 
{\bf any particular reason for the larger radial range for average?
  did we get bigger vortices?:Yes vortices get quite large in quais steady,
 changing the planet off averaging range to [2,10] changes the plots negligibly
 if we want to just use this value range for consistancy } 
% {\bf should state (also for planet-off runs)
% what the radial range is used for averaging the Fourier amplitudes}

@@

In all cases the system remains in a quasi-steady state for
$\gtrsim800P_0$ with a single vortex circulating 
the outer gap edge at the local Keplerian  
frequency. Fig. \ref{Vortex2D} shows a typical plot of the relative 
surface density perturbation in this state. During this stage, the 
vortex intensifies. As the vortex forms it has an initial 
Rossby number $Ro\approx-0.1$ for all $\tilde\beta$
(indicating anti-cyclonic
motion). As the $m=1$ amplitude grows the Rossby number also increases
in magnitude, 
%relative spin of the vortex
%continuously grows along with it
, reaching a characteristic value of $Ro\approx-0.4$ for all cooling
rates. {\bf is the the value reached just before decay for all cooling?:yes 
values peak in small range [-0.35,-0.45] for all beta}

@@

We also measured large increases in the Mach number near
the vortex when it begins to decay. Figure \ref{machplot} plots the
Mach number $M=|\bm{v} - \bm{v}_\mathrm{vor}|/c_s$, where
$\bm{v}_\mathrm{vor}$ corresponds to the bulk velocity of the vortex
around the disc. {\bf check - i assume this is the definition of mach
  number used here:yes this was the definition I used}. Values in Fig. \ref{machplot} have been averaged over
a region within $2H$ of the vortex centre.    
% Mach numbers with repect to the vortex bulk motion around the vortex
% region of $|r-r_{vortex}|<2H$
% are shown in  for various cooling rates over time.
During the quasi-steady state the Mach number increase
steadily, and for all cases $M$ reaches a maximum at 
% with larger growth rate for higher $\tilde\beta$.
a time coincident with the start of vortex decay.

@@

\subsection{Effect of vortex dissipation on the gap structure} 
{\bf are these high resolution results?:yes}
We find vortex growth during the quasi-steady state and its dissipation
also modifies the gap structure. Here, we 
define a gap depth parameter by taking the azimuthal and radial average
of the relative surface density perturbation within $r\in[r_p,r_e]$ where
$r_{e}$ is defined as the outer gap radius such that 
$\langle\Delta \Sigma(r_e)/\Sigma(t=0)\rangle_{\phi}=0$. We are
interested in the region $r \in[r_p,r_e]$ since the vortex is
located at the outer gap edge. A more negative gap depth parameter
correlates with a steeper gap edge. 
{\bf notation clash: rout already used as outer disc boundary - choose
different notation. updated to $r_e$ for now}

@@

\begin{figure}
  \includegraphics[width=\linewidth]{figures/gapchange}
  \caption{Azimuthally averaged gap profiles of relative surface change for
 before and after dissipation for $\tilde\beta=1.0$}
\end{figure}

@@

These results were reproduced in lower resolution simulations as well. Vortex
shocks were similarly seen during vortex dissipation and non-monotonic
 dependence on lifetime with $\tilde\beta$ was similarly found.
Vortex lifetimes are considerably longer in these runs with the maximum
 lifetime longer than the simulation timescale of $2000P_o$ for 
$\tilde\beta=5.0$ which again has $h\approx0.06$.
This inconsistancy in timescales, may be due to resolution as a finer 
grid spacing is required to accurately
reproduce jump conditions along shocks and which cause vortex dissipation.
 {\bf this is the best reason I could personally think of, for lower
   beta runs vortices last slightly longer but for higher values it is
   quite drasticly different. Not sure how numerical dissipation as
   you mentioned would be a factor in this case} 
{\bf you mentioned resolution and convergence to me being a known issue in a previous email}
